---
title: 'Contributors to the 2016 US Presidential Election in California'
author: "Antonio Giannaccari"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  word_document:
    toc: yes
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
```



Campaign donations are important aspects of candidates' presidential races. With  an eye on the upcoming US presidential election of November 2020, let's take a look at some interesting data from the last one in 2016. Specifically, the dataset contains data regarding donors to Democratic and Republican candidates in the State of California, which has historically been a Democrat stronghold. Let's take a look at the data with a view visual tools.




# Unpack the data

This step is important to make the data readily available for the rest of the analysis.
```{r, load_CA_data, warnings= FALSE, message=FALSE}
# Make sure you use vroom() as it is significantly faster than read.csv()
CA_contributors_2016 <- vroom::vroom(here::here("data","CA_contributors_2016.csv"))

zip_url <- "http://www.uszipcodelist.com/download.html"

tables_2 <- zip_url %>% 
  read_html() %>% 
  html_nodes(css="table")

zip_codes <- map(tables_2, . %>% 
               html_table(fill=TRUE)%>% 
               clean_names())
tables_2

```

```{r, load_CA_data2, warnings= FALSE, message=FALSE}
# Make sure you use vroom() as it is significantly faster than read.csv()
CA_contributors_2016 <- vroom::vroom(here::here("data","CA_contributors_2016.csv"))
glimpse(CA_contributors_2016)
```
# Include the two frameworks
```{r CA_contributors_2016_merge}
zip_code_database <-
vroom::vroom(here::here("data","zip_code_database.csv"))

zip_code_database$zip = as.numeric(zip_code_database$zip)

CA_contributors_2016_merge = CA_contributors_2016 %>%
  inner_join(zip_code_database, by = "zip")
glimpse(CA_contributors_2016_merge)
```
# Plot for Hillary Clinton
```{r Clinton_plot}
library(scales)

clinton <- CA_contributors_2016_merge %>%
  filter(cand_nm == "Clinton, Hillary Rodham") %>%
  group_by(primary_city) %>%
  summarise(
    total_contribution = sum(contb_receipt_amt)
  ) %>%
  arrange(desc(total_contribution)) %>%
  slice_max(total_contribution, n=8)

clinton$title <- "Clinton, Hillary Rodham"

p1 <- ggplot(clinton, aes(x = total_contribution, y = reorder(primary_city, total_contribution))) +
  geom_col(fill="steelblue") +
  labs (
    title = "Where did candidates raise most money?",
    x="Amount raised",
    y=NULL
    ) +
  theme_bw() +
  facet_grid(. ~ title) +
  theme(title=element_text(size=8),
        axis.text.x=element_text(size=6),
        axis.text.y=element_text(size=6),
        aspect.ratio = 1) +
  scale_x_continuous(labels = scales::dollar) 

p1

```

# Plot for Donald Trump
```{r Trump_plot}

trump <- CA_contributors_2016_merge %>%
  filter(cand_nm == "Trump, Donald J.") %>%
  group_by(primary_city) %>%
  summarise(
    total_contribution = sum(contb_receipt_amt)
  ) %>%
  arrange(desc(total_contribution)) %>%
  slice_max(total_contribution, n=8)

trump$title <- "Trump, Donald J."

p2 <- ggplot(trump, aes(x = total_contribution, y = reorder(primary_city, total_contribution))) +
  geom_col(fill="red3") +
  labs (
    x="Amount raised",
    y=NULL
    ) +
  theme_bw() +
  facet_grid(. ~ title) +
  theme(title=element_text(size=8),
        axis.text.x=element_text(size=6),
        axis.text.y=element_text(size=6),
        aspect.ratio = 1) +
  scale_x_continuous(labels = scales::dollar) 

p2

```

# Merge the two plots
```{r merge_plot}
library(patchwork)
p1 + p2
```
Unsurprisingly, the largest donations were for Hillary Clinton' campaign. Let's now take a look at  more candidates.

# Top 10 Candidate Plot
```{r merge_10_plot}
library(tidytext)

CA_contributors_2016_merge <- CA_contributors_2016_merge %>%
  group_by(cand_nm, CA_contributors_2016_merge$primary_city) %>%
  mutate(total_contribution_city = sum(contb_receipt_amt))

top_10 <- data.frame(CA_contributors_2016_merge$cand_nm, CA_contributors_2016_merge$primary_city, CA_contributors_2016_merge$total_contribution_city)
top_10 <- unique(top_10)

top_10_name <- CA_contributors_2016_merge %>%
  group_by(cand_nm) %>%
  summarise(total_contribution_all = sum(contb_receipt_amt)) %>%
  arrange(desc(total_contribution_all)) %>%
  slice_max(total_contribution_all, n=8) 

top_10 <- top_10 %>%
  mutate(iftop10 = CA_contributors_2016_merge.cand_nm %in% top_10_name$cand_nm)

top_10 <- top_10[-which(top_10$iftop10 == FALSE),]

top_10_new <- top_10 %>%
  group_by(CA_contributors_2016_merge.cand_nm) %>%
  top_n(-10,desc(CA_contributors_2016_merge.total_contribution_city))

top_10_new %>%
  mutate(CA_contributors_2016_merge.cand_nm=as.factor(CA_contributors_2016_merge.cand_nm), CA_contributors_2016_merge.primary_city=reorder_within(CA_contributors_2016_merge.primary_city, CA_contributors_2016_merge.total_contribution_city, CA_contributors_2016_merge.cand_nm)) %>%
  ggplot(aes(CA_contributors_2016_merge.primary_city, CA_contributors_2016_merge.total_contribution_city, fill = CA_contributors_2016_merge.cand_nm)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~CA_contributors_2016_merge.cand_nm, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "Amount raised",
    x = NULL,
    title = "Where did candidates raise most money?") +
  theme(title=element_text(size=8),
    axis.text.x=element_text(size=5),
    axis.text.y=element_text(size=5))

```
Again, we see that Clinton received by far the largest amount of donations in the State of California. The other Democratic candidate, Bernie Sanders, came as a far second. The remaining 8 candidates are all Republican.







