---
title: 'Drinking Trends around the World'
author: "Antonio Giannaccari"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  word_document:
    toc: yes
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
```



# Where Do People Drink The Most Beer, Wine And Spirits?


```{r, load_alcohol_data}
library(fivethirtyeight)
data(drinks)
```

## Data types and missing values
```{r glimpse_skim_data}
str(drinks)
apply(drinks[,c(2:5)],2,function(x) length(which(x == 0)))
```

When dealing with datasets, it is important to make sure we have a good idea of what the data looks like. As we can see above, the variable types are "character" for one column and "numeric" for the remaining four, with one column  with decimal points. For what concerns missing information, you can see that for some countries such as North Korea no information was recorded regardless of the alcohol type. In other cases, such as Burundi, the information was only partially filled, meaning some types of liquor were recorded. Either way, the missing information would provide sligthly misleading information given the outliers.

## Beer Consumption
```{r beer_plot}
top25beer <- arrange(drinks, desc(drinks$beer_servings)) %>% 
  slice(1:25)
  

print(top25beer)
ggplot(data = top25beer) + aes(x= reorder(country, beer_servings),y=beer_servings) +
  geom_col() +
  ggtitle("Top 25 countries by Beer Consumption") +
  xlab("") +
  ylab("Beer  Servings")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1),
        axis.title.x = element_blank(),
        )+
  coord_flip()+
  NULL
  

```

## Wine Consumption
```{r wine_plot}


top25wine <- arrange(drinks, desc(drinks$wine_servings)) %>% 
  slice(1:25)
  

print(top25wine)
ggplot(data = top25wine) + aes(x= reorder(country, wine_servings),y=wine_servings) +
   ggtitle("Top 25 countries by Wine Consumption") +
  geom_col() +
  xlab("") +
  ylab("Wine  Servings")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1),
        axis.title.x = element_blank(),
        )+
   coord_flip()+
  NULL

```

## Spirit Consumption
```{r spirit_plot}

top25spirit <- arrange(drinks, desc(drinks$spirit_servings)) %>% 
  slice(1:25)
  

print(top25spirit)
ggplot(data = top25spirit) + aes(x= reorder(country, spirit_servings),y=spirit_servings) +
 ggtitle("Top 25 countries by Spirit Consumption") +
  geom_col() +
  xlab("") +
  ylab("Spirit  Servings")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1),
        axis.title.x = element_blank(),
        )+
   coord_flip()+
  NULL
```

The alcohol consumption analysis provides both some expected and some unexpected insights on the drinking trends of countries around  the globe. For what concerns beer, I was not too surprised to see Namibia be the highest consumer: several African countries have displayed excess-drinking trends when it comes  to beer, and with Namibia being a former German colony with strong connections to Germany still to date, there probably are large imports  of beer. Other than Germany, it is not surprising  to see Czech Republic,  Ireland, and Poland in the top 10. 

For what concerns wine, we can notice that many countries that appear on the top 25 are European. I think this is not only the case because wine is a specialty of the region, but also because wine is a more expensive drink compared to beer, therefore it is expected to be consumed in relatively richer countries. Finally, the liquor drinking trends were the ones that displayed more variety, both when it comes to geographic location and economic situation. We see countries in the Caribbean, in Eastern Europe, and Southeast Asia among the largest consumers. This shows that unlike beer and wine, spirits tend to be consumed by a more diverse group of individuals.
