---
title: "IMDB ratings: Differences between directors"
author: "Antonio Giannaccari"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```

#Introduction

As this is the first of my posts, I thought it would make sense to provide some context on the projects I will share. As  part of a team of three students, I analysed data to answer questions and visualise my findings in a clear way. Please find below an example.

# Steve vs Tim
During this project, my teammates and I explored the relationship in the mean IMDB rating for Steven Spielberg and Tim Burton films. We have done this by calculating the confidence intervals for the mean ratings of the two directors and then seeing whether they overlap at the 95% confidence interval. 

# Unpack the data
First of all, lets load the data and examine its structure.

```{r load-movies-data}
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)
```

# Construct the confidence interval
Let's calculate all the necessary metrics to answer this: 

```{r}

movies_cleaned <- movies %>% 
  filter((director=="Tim Burton")|(director=="Steven Spielberg"))

t.test(rating ~ director, data = movies_cleaned)

Steve_vs_Tim <- movies_cleaned %>%
  group_by(director) %>%
  summarise(mean_rating = mean(rating, na.rm = TRUE),
            sd_rating = sd(rating, na.rm=TRUE),
            count = n(),
            se_rating = sd_rating/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_rating,
            lower = mean_rating - t_critical * se_rating,
            upper = mean_rating + t_critical * se_rating
            )

Steve_vs_Tim

```

# Visualise our findings

```{r}

ggplot(Steve_vs_Tim, aes(x=reorder(director, mean_rating))) +
  geom_point(aes(y=mean_rating, color=director, size=4)) +
  geom_rect(aes(xmin=-Inf,xmax=+Inf,ymin=7.275,ymax=7.335),fill="grey",alpha=1)+
  geom_errorbar(aes(ymin=lower, ymax=upper, color=director), width=0.1, size=1.5) +
  coord_flip() +
  labs (
    title = "Do Spielberg and Burton have the same mean in IMDB ratings?",
    y = "Mean IMDB Rating",
    subtitle = "95% confidence intervals overlap"
  ) +
  theme_bw() +
  theme(axis.title.y = element_blank(),
        aspect.ratio = 1/1.5) +
  theme(legend.position = "none")

  
  
```
As you can see, there is some overlap in rating between films by Steven Spielberg and  by Tim Burton, although not too much, at the 95% confidence interval. Ratings for Spielberg tend to be considerably higher than Burton.
